---
title: "Lines and Shapes"
output: html_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lines and shapes

## Polygons and Polylines

Line and polygon data can come from a variety of sources:

-   MULTIPOLYGON, POLYGON, MULTILINESTRING, and LINESTRING objects (from the sf package)
-   map objects (from <maps::map>()); use map(fill = TRUE) for polygons, FALSE for polylines
-   Two-column numeric matrix; the first column is longitude and the second is latitude. Polygons are separated by rows of (NA, NA). It is not possible to represent multi-polygons nor polygons with holes using this method; use sf::st_polygon() instead.

```{r}
library(leaflet)
library(tigris)
```

```{r}
# From https://www.census.gov/geo/maps-data/data/cbf/cbf_state.html
#states <- sf::read_sf("shp/cb_2013_us_state_20m.shp",
#  layer = "cb_2013_us_state_20m")
states <- states()
```

```{r}
names(states)
```

```{r}
ne_states <- subset(states, 
                    states$STUSPS %in% c(
                      "CT", "ME", "VT", "NH", "MA", 
                      "RI", "NY", "NJ", "PA"
                    ))

leaflet(ne_states) |> 
  addPolygons(color = "#444444", weight = 1, 
              smoothFactor = 0.5, opacity = 1.0,
              fillOpacity = 0.5, fillColor = ~colorQuantile(
                "YlOrRd", ALAND
              )(ALAND),
              highlightOptions = highlightOptions(
                color = "white", weight = 2,
                bringToFront = TRUE
              ))
  
```

## Simplifying complex ploygons/polylines

```{r}
full_size <- rnaturalearth::countries110
object.size(full_size)
```

```{r}
simplified <- rmapshaper::ms_simplify(full_size)
object.size(simplified)
```

```{r}
names(simplified)
```

```{r}
plot(simplified$geometry)
```

## Circles

> Requires circle center and radius

```{r}
cities <- read.csv(textConnection("
City,Lat,Long,Pop
Boston,42.3601,-71.0589,645966
Hartford,41.7627,-72.6743,125017
New York City,40.7127,-74.0059,8406000
Philadelphia,39.9500,-75.1667,1553000
Pittsburgh,40.4397,-79.9764,305841
Providence,41.8236,-71.4222,177994
"))
class(cities)
```

```{r}
leaflet(cities) |> addTiles() |> 
  addCircles(lng = ~Long, lat = ~Lat, weight = 1,
             radius = ~sqrt(Pop) * 30, popup = ~City)
```

## Rectangles

```{r}
leaflet() |> addTiles() |> 
  addRectangles(
    lng1=-118.456554, lat1=34.078039,
    lng2=-118.436383, lat2=34.062717,
    fillColor = "transparent"
  )
```
