---
title: "Leaflet Markers"
output: html_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Markers

```{r}
library(leaflet)
```

## Data sources

Point data for markers can come from a variety of sources:

-   `POINT`, `sfc_POINT`, and `sf` objects (from the [sf](https://r-spatial.github.io/sf/) package); only `X` and
    `Y` dimensions will be considered
-   Two-column numeric matrices (first column is longitude, second is
    latitude
-   Data frame with latitude and longitude columns. You can explicitly
    tell the marker function which columns contain the coordinate data
    (e.g., `addMarkers(lng = ~Longitude, lat = ~Latitude)`), or
    let the function look for columns named `lat`/`latitude` and
    `lon`/`lng`/`long`/`longitude` (case insensitive).
-   Simply provide numeric vectors as `lng` and
    `lat` arguments

## Icon markers

```{r}
data("quakes")

leaflet(data = quakes[1:20,]) |> 
  addTiles() |> 
  addMarkers(~long, ~lat, 
             popup = ~as.character(mag),
             label = ~as.character(mag))
```

## Custom icons

> `mkIcon()`

```{r}
greenLeafIcon <- makeIcon(
  iconUrl = "https://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "https://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes[1:4,]) |>  addTiles() |> 
  addMarkers(~long, ~lat, icon = greenLeafIcon)
```

### Multiple icons

```{r}
quakes1 <- quakes[1:10,]

leafIcons <- icons(
  iconUrl = ifelse(quakes1$mag < 4.6,
    "https://leafletjs.com/examples/custom-icons/leaf-green.png",
    "https://leafletjs.com/examples/custom-icons/leaf-red.png"
  ),
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "https://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes1) |> 
  addTiles() |> 
  addMarkers(~long, ~lat, icon = leafIcons)
```

```{r}
# Make a list of icons. We'll index into it based on name.
oceanIcons <- iconList(
  ship = makeIcon(
    "https://upload.wikimedia.org/wikipedia/commons/thumb/0/0b/Maki2-ferry-18.svg/480px-Maki2-ferry-18.svg.png",
    "https://upload.wikimedia.org/wikipedia/commons/thumb/0/0b/Maki2-ferry-18.svg/18px-Maki2-ferry-18.svg.png",
    18,
    18
  ),
  pirate = makeIcon(
    "https://upload.wikimedia.org/wikipedia/commons/thumb/c/c3/Maki2-danger-24.svg/240px-Maki2-danger-24.svg.png",
    "https://upload.wikimedia.org/wikipedia/commons/thumb/c/c3/Maki2-danger-24.svg/24px-Maki2-danger-24.svg.png",
    24,
    24
  )
)

# Some fake data
df <- sf::st_as_sf(
  data.frame(
    type = factor(ifelse(runif(20) > 0.75, "pirate", "ship"), c("ship", "pirate")),
    lng = (runif(20) - .5) * 10 - 90.620130,
    lat = (runif(20) - .5) * 3.8 + 25.638077
  ),
  coords = c("lng", "lat"),
  crs = 4326
)

leaflet(df) %>% addTiles() %>%
  # Select from oceanIcons based on df$type
  addMarkers(icon = ~oceanIcons[type])
```

## Awesome icons

> -   The `library` argument has to be one of ‘ion’, ‘fa’, or
>     ‘glyphicon’
> -   The `icon` argument needs to be the name of any
>     valid icon supported by the the respective library (w/o the prefix of
>     the library name)

```{r}
df.20 <- quakes[1:20,]

getColor <- function(quakes) {
  sapply(quakes$mag, function(mag) {
    if(mag <= 4) {
      "green"
    } else if (mag <= 5) {
      "orange"
    } else {
      "red"
    }
  })
}

icons <- awesomeIcons(
  icon = "ios-close", 
  iconColor = "black", 
  library = "ion",
  markerColor = getColor(df.20)
)

df.20 |> leaflet() |> addTiles() |> 
  addAwesomeMarkers(~long, ~lat, icon=icons, 
                    label = ~as.character(mag))
```

## Marker clusters

```{r}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions()
)
```

## Circle markers

> Constant radius regardless of zoom level
>
> Default

```{r}
leaflet(df) |> addTiles() |>  addCircleMarkers()
```

> Customized

```{r}
pal <- colorFactor(c("navy", "red"), domain = c("ship", "pirate"))

leaflet(df) |> addTiles() |> 
  addCircleMarkers(
    radius = ~ifelse(type == "ship", 6, 10),
    color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5
  )
```
