---
title: "Leaflet Tutorial"
output: html_document
date: "2025-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Leaflet map widget

```{r}
library(leaflet)
```

> Create with options

```{r}
l <- leaflet(options = leafletOptions(minZoom = 0, maxZoom = 18))
```

-   [`setView()`](https://rstudio.github.io/leaflet/reference/map-methods.html) sets the center of the map view and the zoom level;
-   [`fitBounds()`](https://rstudio.github.io/leaflet/reference/map-methods.html) fits the view into the rectangle `[lng1, lat1]` – `[lng2, lat2]`;
-   [`clearBounds()`](https://rstudio.github.io/leaflet/reference/map-methods.html) clears the bound, so that the view will be automatically determined by the range of latitude/longitude data in the map layers if provided;

## The Data Object

Accepts:

-   lng/lat matrix
-   data frame with lng/lat columns
-   data frame from `maps`
-   simple features from `sf`

Accepts names `lat`, `latitude`,`lng`, `long`, `longitude`

```{r}
df <- data.frame(lat = 1:10, long = rnorm(10))
leaflet(df) |> addCircles()
```

> Explicit

```{r}
leaflet(df) |> addCircles(lng = ~long, lat = ~lat) 
```

```{r}
library(sf)
polygon1 <- st_polygon(list(cbind(c(2, 4, 4, 1, 2), c(2, 3, 5, 4, 2))))
polygon2 <- st_polygon(list(cbind(c(5, 4, 2, 5), c(2, 3, 2, 2))))
polygon3 <- st_polygon(list(cbind(c(4, 4, 5, 10, 4), c(5, 3, 2, 5, 5))))
polygon4 <- st_polygon(list(cbind(c(5, 6, 6, 5, 5), c(4, 4, 3, 3, 4))))
multi_polygon <- st_multipolygon(list(
    list(cbind(c(4, 4, 5, 10, 4), c(5, 3, 2, 5, 5))),
    list(cbind(c(5, 6, 6, 5, 5), c(4, 4, 3, 3, 4)))
))
mp_test <- st_multipolygon(list(polygon3, polygon4))
mp_test == multi_polygon
multi_polygon
sf_polygons <- st_sf(geometry = st_sfc(polygon1, polygon2, multi_polygon))
leaflet(height = "300px") %>% addPolygons(data = sf_polygons)
```

```{r}
mp_test
```

```{r}
identical(multi_polygon$geometry, mp_test$geometry)
```

```{r}
library(maps)
mapStates = map("state", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) |> 
  addTiles() |> 
  addPolygons(fillColor = topo.colors(10, alpha = NULL),
              stroke = FALSE)
```

## Formula interface

```{r}
df <- data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m <- leaflet(df) |> addTiles()
m |> addCircleMarkers(radius = ~size, color = ~color, 
                      fill = FALSE)
```

```{r}
m |> addCircleMarkers(radius = runif(100, 4, 10),
                      color = c('red'))
```
