---
title: "LeafletColors"
output: html_document
date: "2025-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Leaflet Colors

```{r}
library(leaflet)
```

The `color*()` functions generate *palette* functions which accept a vector of input values and returns a vector of colors.

-   `colorNumeric()`
-   `colorBin()`
-   `colorQuantile()`
-   `colorFactor()` \<- for categorical data

```{r}
pal <- colorNumeric(c("red", "green", "blue"), 1:10)
pal(c(1,6,9))
```

The `palette` argument specifies the colors to map the
data to. This argument can take one of several forms:

1.  The name of a preset palette from the `RColorBrewer`
    package, e.g., `"RdYlBu"`, `"Accent"`, or
    `"Greens"`.
2.  The full name of a `viridis` palette:
    `"magma"`, `"inferno"`, `"plasma"`,
    `"viridis"`, `"cividis"`, `"rocket"`,
    `"mako"`, or `"turbo"`.
3.  A character vector of RGB or named colors, e.g.,
    [`palette()`](https://rdrr.io/r/grDevices/palette.html), `c("#000000", "#0000FF", "#FFFFFF")`,
    `topo.colors(10)`.
4.  A function that receives a single value between 0 and 1 and returns
    a color, e.g.,:
    `colorRamp(c("#000000", "#FFFFFF"), interpolate="spline")`

If `domain` is set to `NULL`, the palette is generated each time. This will provide inconsistent coloring/scaling.

## Continuous data

```{r}
library(sf)
countries <- read_sf("https://rstudio.github.io/leaflet/json/countries.geojson")
map <- leaflet(countries)
```

```{r}
countries[1,]
```

```{r}
hist(countries$gdp_md_est, breaks = 20, main = "")
```

```{r}
par(mar = c(5,5,0,0), cex = 0.8)
hist(countries$gdp_md_est, breaks = 20, main = "")
```

### Continuous input, continuous colors

> `colorNumeric`

While the "Blues" palette only contains nine colors, it interpolates them providing continuous colors.

```{r}
pal <- colorNumeric(
  palette = "Blues", 
  domain = countries$gdp_md_est
)
pal1 <- colorNumeric(
  palette = c("white", "navy"), 
  domain = countries$gdp_md_est
)
pal2 <- colorNumeric(
  palette = c("#FFFFFF","#0000FF", "#000088"), 
  domain = countries$gdp_md_est
)
```

```{r}
map |> 
  addPolygons(stroke = FALSE, smoothFactor = 0.2,
              fillOpacity = 1,
              color = ~pal2(gdp_md_est))
```

### Continuous input, discrete colors

> `colorBin()`
>
> Can specify number of bins or exact breakes.

```{r}
binpal <- colorBin("Blues", countries$gdp_md_est,
                   6, pretty = FALSE)
map |> addPolygons(
  stroke = FALSE, 
  smoothFactor = 0.2,
  fillOpacity = 1,
  color = ~binpal(gdp_md_est) 
)
```

> `colorQuantile()` - equal number of observations

```{r}
qpal <- colorQuantile("Blues", countries$gdp_md_est,
                      n = 7)
map %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~qpal(gdp_md_est))
```

## Categorical data

If there are more factor levels than palette elements, the palette colors will be interpolated.

Accepts either a factor or character vector, or the `levels` parameter.

```{r}
countries$category <- factor(sample.int(5L, nrow(countries), TRUE))

factpal <- colorFactor(topo.colors(5), 
                       countries$category)
leaflet(countries) |> 
  addPolygons(stroke = F, smoothFactor = 0.2,
              fillOpacity = 1, color = ~factpal(category))
```
